---
title: "Predict Student Romantic Relationship Status"
author: "Weiyang Sheng"
format: pdf
editor: visual
---

```{r}
#| echo: false
#| message: false
#| warning: false

## keep this chunk!
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  tidy = "styler",
  fig.align = "center"
)
```

## Introduction

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidymodels)
library(vip)
library(DALEX)
library(stacks)
options(yardstick.event_first = FALSE)
library(discrim)
library(xgboost)
library(kernlab)
library(nnet)
library(klaR)
set.seed(123)
```

```{r import}
student_por <- read_csv("student-por.csv")
student_mat <- read_csv("student-mat.csv")

student_mat <- student_mat %>% mutate(course = "Math")
student_por <- student_por %>% mutate(course = "Portuguese")
student_all <- bind_rows(student_mat, student_por)

```

```{r}
students <- student_all %>%
  mutate(
    romantic   = factor(romantic, levels = c("no", "yes")),
    sex        = factor(sex),
    address    = factor(address),
    Mjob       = factor(Mjob),
    Fjob       = factor(Fjob),
   Medu       = as.numeric(Medu),
Fedu       = as.numeric(Fedu),
Medu_pct   = 100 * Medu / 4,
Fedu_pct   = 100 * Fedu / 4,
edu_gap_pct = abs(Medu_pct - Fedu_pct),
    Mjob_ses = case_when(
      Mjob %in% c("teacher", "health") ~ "higher",
      Mjob %in% c("services") ~ "middle",
      TRUE ~ "lower"
    ),
    Mjob_ses = factor(Mjob_ses, levels = c("lower","middle","higher")),
    Fjob_ses = case_when(
      Fjob %in% c("teacher", "health") ~ "higher",
      Fjob %in% c("services") ~ "middle",
      TRUE ~ "lower"
    ),
    Fjob_ses = factor(Fjob_ses, levels = c("lower","middle","higher")),
    activities = factor(activities, levels = c("no","yes")),
    higher     = factor(higher,     levels = c("no","yes")),
    internet   = factor(internet,   levels = c("no","yes")),
    age        = as.numeric(age),
    studytime  = as.numeric(studytime),
    freetime   = as.numeric(freetime),
    goout      = as.numeric(goout),
    Dalc       = as.numeric(Dalc),
    Walc       = as.numeric(Walc),
    age_group = case_when(
      age <= 16 ~ "15-16",
      age == 17 ~ "17",
      age == 18 ~ "18",
      age >= 19 ~ "19+"
    ),
    age_group = factor(age_group, levels = c("15-16","17","18","19+")),
    alc_mean = (Dalc + Walc) / 2,
    alc_risk = factor(if_else(Walc >= 4 | Dalc >= 3, "higher", "lower"),
                      levels = c("lower","higher")),
    freetime_pct  = 100 * (freetime  - 1) / 4,
    goout_pct     = 100 * (goout     - 1) / 4,
    studytime_pct = 100 * (studytime - 1) / 3,
    alc_mean_pct  = 100 * (alc_mean  - 1) / 4
  )

set.seed(123)
splits <- initial_split(students, prop = 0.8, strata = romantic)
train_data <- training(splits)
test_data  <- testing(splits)
train_data_mod <- train_data %>% dplyr::select(-sex)
test_data_mod  <- test_data  %>% dplyr::select(-sex)
train_F <- train_data %>% filter(sex == "F")
test_F  <- test_data  %>% filter(sex == "F")

train_M <- train_data %>% filter(sex == "M")
test_M  <- test_data  %>% filter(sex == "M")
train_F_mod <- train_F %>% dplyr::select(-sex)
test_F_mod  <- test_F  %>% dplyr::select(-sex)

train_M_mod <- train_M %>% dplyr::select(-sex)
test_M_mod  <- test_M  %>% dplyr::select(-sex)

train_M_mod <- train_M %>% dplyr::select(-sex)
test_M_mod  <- test_M  %>% dplyr::select(-sex)

prop.table(table(train_data$romantic))
prop.table(table(test_data$romantic))
```

## Exploratory Analysis

```{r}
edu_base <- students %>%
  group_by(sex) %>%
  summarise(base = mean(romantic == "yes"), .groups = "drop")

edu_parent <- students %>%
  dplyr::select(sex, romantic, Medu_pct, Fedu_pct) %>%
  pivot_longer(
    cols = c(Medu_pct, Fedu_pct),
    names_to = "parent",
    values_to = "edu_pct"
  ) %>%
  mutate(
    parent = recode(parent, Medu_pct = "Mother", Fedu_pct = "Father"),
    edu_pct = factor(edu_pct, levels = c(0, 25, 50, 75, 100))
  ) %>%
  group_by(sex, parent, edu_pct) %>%
  summarise(prop_yes = mean(romantic == "yes"), n = n(), .groups = "drop") %>%
  left_join(edu_base, by = "sex") %>%
  mutate(delta_pp = 100 * (prop_yes - base))

pad_parent <- 2

edu_parent_mother <- edu_parent %>% filter(parent == "Mother")
ylim_m <- range(edu_parent_mother$delta_pp, na.rm = TRUE)

ggplot(edu_parent_mother, aes(x = edu_pct, y = delta_pp, group = 1)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_line(linewidth = 1.2) +
  geom_point(aes(size = n), alpha = 0.9) +
  facet_wrap(~ sex, nrow = 1) +
  scale_y_continuous(labels = function(x) paste0(x, " pp")) +
  coord_cartesian(ylim = c(ylim_m[1] - pad_parent, ylim_m[2] + pad_parent)) +
  labs(
    title = "Mother education and relationship rate (difference from sex average)",
    x = "Education level (0–100%)",
    y = "Difference from sex average (percentage points)",
    size = "N"
  ) +
  theme_minimal()

edu_parent_father <- edu_parent %>% filter(parent == "Father")
ylim_f <- range(edu_parent_father$delta_pp, na.rm = TRUE)

ggplot(edu_parent_father, aes(x = edu_pct, y = delta_pp, group = 1)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_line(linewidth = 1.2) +
  geom_point(aes(size = n), alpha = 0.9) +
  facet_wrap(~ sex, nrow = 1) +
  scale_y_continuous(labels = function(x) paste0(x, " pp")) +
  coord_cartesian(ylim = c(ylim_f[1] - pad_parent, ylim_f[2] + pad_parent)) +
  labs(
    title = "Father education and relationship rate (difference from sex average)",
    x = "Education level (0–100%)",
    y = "Difference from sex average (percentage points)",
    size = "N"
  ) +
  theme_minimal()

```

```{r}
# ===== Dependency check example: does parental SES predict goout? =====
goout_lm <- lm(goout ~ Medu_pct + Fedu_pct + Mjob_ses + Fjob_ses + sex, data = students)
summary(goout_lm)

# A simple visualization (mean goout by parental SES)
students %>%
  group_by(sex, Mjob_ses) %>%
  summarise(goout_mean = mean(goout, na.rm = TRUE), n = n(), .groups = "drop") %>%
  ggplot(aes(x = Mjob_ses, y = goout_mean)) +
  geom_col() +
  facet_wrap(~ sex) +
  labs(
    title = "Is going-out behavior associated with mother's job SES?",
    x = "Mother job SES",
    y = "Mean goout (1–5)"
  ) +
  theme_minimal()
```

```{r}
students %>%
  group_by(sex, edu_gap_pct) %>%
  summarise(prop_yes = mean(romantic == "yes"), .groups = "drop") %>%
  ggplot(aes(x = edu_gap_pct, y = prop_yes, color = sex, group = sex)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 2.5) +
  scale_x_continuous(breaks = c(0, 25, 50, 75, 100)) +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  labs(
    title = "Education Gap (0–100%) vs Relationship Rate",
    x = "Gap between parents' education (0–100%)",
    y = "Proportion (romantic = yes)",
    color = "Sex"
  ) +
  theme_minimal()
```

```{r}
students %>%
  group_by(sex, edu_gap_pct) %>%
  summarise(
    n = n(),
    yes_n = sum(romantic == "yes"),
    prop_yes = mean(romantic == "yes"),
    .groups = "drop"
  ) %>%
  arrange(sex, edu_gap_pct)
```

```{r}
gap_base <- students %>%
  group_by(sex) %>%
  summarise(base = mean(romantic == "yes"), .groups = "drop")

gap_df <- students %>%
  mutate(
    gap_group = case_when(
      edu_gap_pct == 0 ~ "0",
      edu_gap_pct == 25 ~ "25",
      edu_gap_pct == 50 ~ "50",
      edu_gap_pct >= 75 ~ "75-100"
    ),
    gap_group = factor(gap_group, levels = c("0", "25", "50", "75-100"))
  ) %>%
  group_by(sex, gap_group) %>%
  summarise(prop_yes = mean(romantic == "yes"), n = n(), .groups = "drop") %>%
  left_join(gap_base, by = "sex") %>%
  mutate(delta_pp = 100 * (prop_yes - base))

ylim_gap <- range(gap_df$delta_pp, na.rm = TRUE)
pad_gap <- 2

ggplot(gap_df, aes(x = gap_group, y = delta_pp, group = sex, color = sex)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_line(linewidth = 1.2) +
  geom_point(aes(size = n), alpha = 0.9) +
  scale_y_continuous(labels = function(x) paste0(x, " pp")) +
  coord_cartesian(ylim = c(ylim_gap[1] - pad_gap, ylim_gap[2] + pad_gap)) +
  labs(
    title = "Education gap and relationship rate (difference from sex average)",
    x = "Gap between parents' education",
    y = "Difference from sex average (percentage points)",
    color = "Sex",
    size = "N"
  ) +
  theme_minimal()
```

```{r}
edu_combo <- students %>%
  mutate(
    Medu_cat = factor(Medu, levels = 0:4),
    Fedu_cat = factor(Fedu, levels = 0:4)
  ) %>%
  group_by(sex, Medu_cat, Fedu_cat) %>%
  summarise(
    yes_n = sum(romantic == "yes"),
    n = n(),
    prop_yes = yes_n / n,
    .groups = "drop"
  ) %>%
  mutate(
    label = paste0("yes=", yes_n, "\n", scales::percent(prop_yes, accuracy = 1)),
   
    text_col = if_else(yes_n >= median(yes_n, na.rm = TRUE), "white", "black")
  )

fill_max <- max(edu_combo$yes_n, na.rm = TRUE)

plot_heatmap_yesn <- function(df_sex, sex_label) {
  ggplot(df_sex, aes(x = Medu_cat, y = Fedu_cat, fill = yes_n)) +
    geom_tile(color = "white", linewidth = 0.5) +
    geom_text(aes(label = label, color = text_col), size = 3) +
    scale_color_identity() +
    scale_fill_gradient(
      low = "white",
      high = "orange",
      limits = c(0, fill_max),
      breaks = scales::pretty_breaks(n = 5),
      na.value = "grey90"
    ) +
    labs(
      title = paste0("Count of students in a relationship by parents' education (", sex_label, ")"),
      x = "Mother education (0–4)",
      y = "Father education (0–4)",
      fill = "Count (romantic = yes)"
    ) +
    theme_minimal()
}

plot_heatmap_yesn(edu_combo %>% filter(sex == "F"), "Female")
plot_heatmap_yesn(edu_combo %>% filter(sex == "M"), "Male")
```

```{r}
students %>%
  group_by(sex) %>%
  summarise(prop_yes = mean(romantic == "yes"), n = n(), .groups = "drop") %>%
  ggplot(aes(x = sex, y = prop_yes)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  labs(
    title = "Relationship Rate by Sex",
    x = "Sex",
    y = "Proportion (romantic = yes)"
  ) +
  theme_minimal()

```

```{r}
students %>%
  dplyr::select(sex, romantic, Mjob_ses, Fjob_ses) %>%
  pivot_longer(cols = c(Mjob_ses, Fjob_ses), names_to = "parent", values_to = "job_ses") %>%
  mutate(parent = recode(parent, Mjob_ses = "Mother", Fjob_ses = "Father")) %>%
  group_by(sex, parent, job_ses) %>%
  summarise(prop_yes = mean(romantic == "yes"), .groups = "drop") %>%
  ggplot(aes(x = job_ses, y = prop_yes)) +
  geom_col() +
  facet_grid(sex ~ parent) +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  labs(
    title = " Parental Job SES vs Relationship Rate (by Sex)",
    x = "Job SES",
    y = "Proportion (romantic = yes)"
  ) +
  theme_minimal()

```

```{r}
students %>%
  group_by(sex, freetime_pct) %>%
  summarise(prop_yes = mean(romantic == "yes"), .groups = "drop") %>%
  ggplot(aes(x = freetime_pct, y = prop_yes, color = sex, group = sex)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3) +
  scale_x_continuous(breaks = c(0, 25, 50, 75, 100)) +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  labs(
    title = "Free-time (0–100%) × Sex",
    x = "Free-time (0–100%)",
    y = "Proportion (romantic = yes)",
    color = "Sex"
  ) +
  theme_minimal()
```

```{r}
students %>%
  group_by(sex, alc_risk) %>%
  summarise(prop_yes = mean(romantic == "yes"), .groups = "drop") %>%
  ggplot(aes(x = alc_risk, y = prop_yes)) +
  geom_col() +
  facet_wrap(~ sex) +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  labs(
    title = "Alcohol Risk vs Relationship Rate (by Sex)",
    x = "Alcohol risk",
    y = "Proportion (romantic = yes)"
  ) +
  theme_minimal()
```

## Modeling Approach

```{r}
rec_F <- recipe(romantic ~ ., data = train_F_mod) %>%
  step_rm(age, Dalc, Walc, Mjob, Fjob, Medu, Fedu,
          freetime_pct, goout_pct, studytime_pct, alc_mean_pct) %>%
  step_normalize(all_numeric_predictors()) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_zv(all_predictors())

rec_M <- recipe(romantic ~ ., data = train_M_mod) %>%
  step_rm(age, Dalc, Walc, Mjob, Fjob, Medu, Fedu,
          freetime_pct, goout_pct, studytime_pct, alc_mean_pct) %>%
  step_normalize(all_numeric_predictors()) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_zv(all_predictors())

rec_all <- recipe(romantic ~ ., data = train_data_mod) %>%
  step_rm(age, Dalc, Walc, Mjob, Fjob, Medu, Fedu,
          freetime_pct, goout_pct, studytime_pct, alc_mean_pct) %>%
  step_normalize(all_numeric_predictors()) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_zv(all_predictors())
```

## Logistic Regression

```{r}
logit_spec <- logistic_reg() %>% set_engine("glm")

logit_wf <- workflow() %>%
  add_recipe(rec_all) %>%
  add_model(logit_spec)
logit_fit <- fit(logit_wf, data = train_data_mod)

logit_wf_F <- workflow() %>%
  add_recipe(rec_F) %>%
  add_model(logit_spec)
logit_fit_F <- fit(logit_wf_F, data = train_F_mod)

logit_wf_M <- workflow() %>%
  add_recipe(rec_M) %>%
  add_model(logit_spec)
logit_fit_M <- fit(logit_wf_M, data = train_M_mod)

logit_fit %>%
  extract_fit_parsnip() %>%
  tidy() %>%
  arrange(p.value)

logit_fit_F %>%
  extract_fit_parsnip() %>%
  tidy() %>%
  arrange(p.value)

logit_fit_M %>%
  extract_fit_parsnip() %>%
  tidy() %>%
  arrange(p.value)
```

```{r}
metric_auc <- metric_set(roc_auc)

# (Optional) speed-up: run resamples in parallel
# install.packages("doParallel") if needed
# library(doParallel)
# registerDoParallel(cores = parallel::detectCores(logical = FALSE))

# ============ 1) Random Forest (tuned) ============
rf_spec_tune <- rand_forest(
  mode  = "classification",
  trees = tune(),
  mtry  = tune(),
  min_n = tune()
) %>%
  set_engine("ranger", importance = "impurity")

rf_wf_tune <- workflow() %>%
  add_formula(romantic ~ age_group + address + studytime + freetime + goout +
                activities + alc_mean + higher + internet +
                Medu_pct + Fedu_pct + edu_gap_pct + Mjob_ses + Fjob_ses) %>%
  add_model(rf_spec_tune)

rf_params <- hardhat::extract_parameter_set_dials(rf_spec_tune) %>%
  update(
    trees = dials::trees(c(400, 1600)),
    mtry  = dials::mtry(c(2, 14)),
    min_n = dials::min_n(c(2, 25))
  )

# Smaller grids to reduce runtime (space-filling, not brute force)
rf_grid_small <- dials::grid_space_filling(rf_params, size = 8)   # for choosing v
rf_grid       <- dials::grid_space_filling(rf_params, size = 15)  # final tuning

# ============ 2) Other models (tuned, smaller search) ============
enet_spec <- logistic_reg(penalty = tune(), mixture = tune()) %>%
  set_engine("glmnet", maxit = 200000) %>%
  set_mode("classification")

svm_spec <- svm_rbf(cost = tune(), rbf_sigma = tune()) %>%
  set_engine("kernlab") %>%
  set_mode("classification")

# kNN: tune only the most important hyperparams (faster)
knn_spec <- nearest_neighbor(neighbors = tune(), weight_func = tune()) %>%
  set_engine("kknn") %>%
  set_mode("classification")

mlp_spec <- mlp(hidden_units = tune(), penalty = tune(), epochs = tune()) %>%
  set_engine("nnet") %>%
  set_mode("classification")

# XGBoost: reduce tuned dimensions to avoid slow brute force
xgb_spec <- boost_tree(
  trees       = tune(),
  tree_depth  = tune(),
  learn_rate  = tune(),
  mtry        = tune(),
  min_n       = tune(),
  sample_size = tune()
) %>%
  set_engine("xgboost", verbosity = 0) %>%
  set_mode("classification")

# Decision tree: tune hyperparameters (requested in Progress report 2)
tree_spec_tune <- decision_tree(
  cost_complexity = tune(),
  tree_depth      = tune(),
  min_n           = tune()
) %>%
  set_engine("rpart") %>%
  set_mode("classification")

tree_wf_tune <- workflow() %>%
  add_formula(romantic ~ age_group + address + studytime + freetime + goout +
                activities + alc_mean + higher + internet +
                Medu_pct + Fedu_pct + edu_gap_pct + Mjob_ses + Fjob_ses) %>%
  add_model(tree_spec_tune)

# ---- helper: n predictors after recipe (for xgb mtry range)
get_n_pred <- function(train_df, rec_obj) {
  rec_obj %>%
    prep(training = train_df, retain = TRUE) %>%
    juice() %>%
    dplyr::select(-romantic) %>%
    ncol()
}

# ---- choose best v for RF quickly
choose_best_v_rf <- function(train_df, v_candidates = c(5, 10)) {
  v_tbl <- purrr::map_dfr(v_candidates, function(vv) {
    set.seed(123)
    folds <- vfold_cv(train_df, v = vv, strata = romantic)

    set.seed(123)
    res <- tune_grid(
      rf_wf_tune,
      resamples = folds,
      grid = rf_grid_small,
      metrics = metric_auc,
      control = tune::control_grid(save_pred = FALSE)
    )

    best_auc <- show_best(res, metric = "roc_auc", n = 1) %>% dplyr::pull(mean)
    tibble(v = vv, best_auc = best_auc)
  })

  best_v <- v_tbl %>%
    slice_max(best_auc, n = 1, with_ties = FALSE) %>%
    pull(v)

  list(best_v = best_v, v_tbl = v_tbl)
}

fit_one_group_ensemble <- function(train_df, rec_obj, v_candidates = c(5, 10)) {

  # 1) choose v
  v_obj <- choose_best_v_rf(train_df, v_candidates = v_candidates)
  set.seed(123)
  folds <- vfold_cv(train_df, v = v_obj$best_v, strata = romantic)

  # 2) workflows (recipe models)
  enet_wf <- workflow() %>% add_recipe(rec_obj) %>% add_model(enet_spec)
  svm_wf  <- workflow() %>% add_recipe(rec_obj) %>% add_model(svm_spec)
  knn_wf  <- workflow() %>% add_recipe(rec_obj) %>% add_model(knn_spec)
  mlp_wf  <- workflow() %>% add_recipe(rec_obj) %>% add_model(mlp_spec)
  xgb_wf  <- workflow() %>% add_recipe(rec_obj) %>% add_model(xgb_spec)

  # 3) small, space-filling grids
  enet_grid <- dials::grid_space_filling(hardhat::extract_parameter_set_dials(enet_spec), size = 12)
  svm_grid  <- dials::grid_space_filling(hardhat::extract_parameter_set_dials(svm_spec),  size = 12)

  knn_params <- hardhat::extract_parameter_set_dials(knn_spec) %>%
    update(neighbors = dials::neighbors(c(3, 60)))
  knn_grid <- dials::grid_space_filling(knn_params, size = 15)

  mlp_grid <- dials::grid_space_filling(hardhat::extract_parameter_set_dials(mlp_spec), size = 12)

  n_pred <- get_n_pred(train_df, rec_obj)
  xgb_params <- hardhat::extract_parameter_set_dials(xgb_spec) %>%
    update(
      mtry       = dials::mtry(c(2, n_pred)),
      trees      = dials::trees(c(300, 1500)),
      tree_depth = dials::tree_depth(c(2, 8)),
      min_n      = dials::min_n(c(2, 25)),
      learn_rate = dials::learn_rate(c(-4, -1)),      # 10^-4 to 10^-1
      sample_size = dials::sample_prop(c(0.6, 1.0))
    )
  xgb_grid <- dials::grid_space_filling(xgb_params, size = 15)

  tree_params <- hardhat::extract_parameter_set_dials(tree_spec_tune) %>%
    update(
      tree_depth      = dials::tree_depth(c(1, 10)),
      min_n           = dials::min_n(c(2, 30)),
      cost_complexity = dials::cost_complexity(c(-4, -1))
    )
  tree_grid <- dials::grid_space_filling(tree_params, size = 15)

  # Stack-compatible control: save predictions + workflow
  ctrl_stack <- tune::control_grid(save_pred = TRUE, save_workflow = TRUE)

  # 4) tune
  set.seed(123)
  rf_res   <- tune_grid(rf_wf_tune,   resamples = folds, grid = rf_grid,
                        metrics = metric_auc, control = ctrl_stack)

  set.seed(123)
  enet_res <- tune_grid(enet_wf, resamples = folds, grid = enet_grid,
                        metrics = metric_auc, control = ctrl_stack)

  set.seed(123)
  svm_res  <- tune_grid(svm_wf,  resamples = folds, grid = svm_grid,
                        metrics = metric_auc, control = ctrl_stack)

  set.seed(123)
  knn_res  <- tune_grid(knn_wf,  resamples = folds, grid = knn_grid,
                        metrics = metric_auc, control = ctrl_stack)

  set.seed(123)
  mlp_res  <- tune_grid(mlp_wf,  resamples = folds, grid = mlp_grid,
                        metrics = metric_auc, control = ctrl_stack)

  set.seed(123)
  xgb_res  <- tune_grid(xgb_wf,  resamples = folds, grid = xgb_grid,
                        metrics = metric_auc, control = ctrl_stack)

  set.seed(123)
  tree_res <- tune_grid(tree_wf_tune, resamples = folds, grid = tree_grid,
                        metrics = metric_auc, control = ctrl_stack)

  # 5) stacking
  stack_fit <- stacks() %>%
    add_candidates(rf_res) %>%
    add_candidates(enet_res) %>%
    add_candidates(svm_res) %>%
    add_candidates(knn_res) %>%
    add_candidates(mlp_res) %>%
    add_candidates(xgb_res) %>%
    add_candidates(tree_res) %>%
    blend_predictions(metric = metric_auc) %>%
    fit_members()

  # 6) fit best models for final evaluation on test
  fit_best <- function(wf_obj, res_obj, train_df_inner) {
    best_params <- tune::select_best(res_obj, metric = "roc_auc")
    wf_final <- tune::finalize_workflow(wf_obj, best_params)
    fit(wf_final, data = train_df_inner)
  }

  list(
    best_v = v_obj$best_v,
    v_compare = v_obj$v_tbl,

    rf_best   = tune::select_best(rf_res,   metric = "roc_auc"),
    knn_best  = tune::select_best(knn_res,  metric = "roc_auc"),
    xgb_best  = tune::select_best(xgb_res,  metric = "roc_auc"),
    tree_best = tune::select_best(tree_res, metric = "roc_auc"),

    rf_fit   = fit_best(rf_wf_tune, rf_res, train_df),
    enet_fit = fit_best(enet_wf, enet_res, train_df),
    svm_fit  = fit_best(svm_wf,  svm_res,  train_df),
    knn_fit  = fit_best(knn_wf,  knn_res,  train_df),
    mlp_fit  = fit_best(mlp_wf,  mlp_res,  train_df),
    xgb_fit  = fit_best(xgb_wf,  xgb_res,  train_df),
    tree_fit = fit_best(tree_wf_tune, tree_res, train_df),

    stack_fit = stack_fit,

    # keep tuning results for later diagnostics
    knn_res = knn_res,
    xgb_res = xgb_res,
    rf_res  = rf_res,
    tree_res = tree_res
  )
}

ens_all <- fit_one_group_ensemble(train_data_mod, rec_all, v_candidates = c(5, 10))
ens_F   <- fit_one_group_ensemble(train_F_mod,   rec_F,   v_candidates = c(5, 10))
ens_M   <- fit_one_group_ensemble(train_M_mod,   rec_M,   v_candidates = c(5, 10))

rf_fit       <- ens_all$rf_fit
rf_fit_F     <- ens_F$rf_fit
rf_fit_M     <- ens_M$rf_fit

enet_fit     <- ens_all$enet_fit
svm_fit      <- ens_all$svm_fit
knn_fit      <- ens_all$knn_fit
mlp_fit      <- ens_all$mlp_fit
xgb_fit      <- ens_all$xgb_fit
tree_fit     <- ens_all$tree_fit

enet_fit_F   <- ens_F$enet_fit
svm_fit_F    <- ens_F$svm_fit
knn_fit_F    <- ens_F$knn_fit
mlp_fit_F    <- ens_F$mlp_fit
xgb_fit_F    <- ens_F$xgb_fit
tree_fit_F   <- ens_F$tree_fit

enet_fit_M   <- ens_M$enet_fit
svm_fit_M    <- ens_M$svm_fit
knn_fit_M    <- ens_M$knn_fit
mlp_fit_M    <- ens_M$mlp_fit
xgb_fit_M    <- ens_M$xgb_fit
tree_fit_M   <- ens_M$tree_fit

stack_fit_all <- ens_all$stack_fit
stack_fit_F   <- ens_F$stack_fit
stack_fit_M   <- ens_M$stack_fit

# For reporting:
ens_all$best_v
ens_all$rf_best
ens_all$knn_best
ens_all$xgb_best
ens_all$tree_best
```

Finally, according to the results of random forest analysis, as shown in the plot, age, frequency of going out, free time, drinking, parents' family background, and study time show significantly higher importance than other variables. Among them, students who are older and have a higher frequency of going out are most likely to be in a relationship. After this, I will do the model evaluation to dicide which model result should I finally pick. I also conducted 5-fold cross-validation on the training set. The average cross-validated accuracy was 78.4%, F1 score was 0.84, and AUC was 0.86 (they might be different when I render). These results show that random forest performs well, even when trained on different data subsets, making it a reliable choice for this prediction task.

## Model Performance Comparison

Now I compare how the three models performed on the test set (209 students who were not used in training). I consider each model’s ability to correctly classify students and to rank them by probability of being in love.

```{r}
best_f1_threshold <- function(truth, prob, grid = seq(0.05, 0.95, by = 0.01)) {
  truth <- factor(truth, levels = c("no", "yes"))

  tibble(threshold = grid) %>%
    mutate(
      f1 = purrr::map_dbl(threshold, function(t) {
        pred <- factor(dplyr::if_else(prob >= t, "yes", "no"),
                       levels = c("no", "yes"))
        yardstick::f_meas_vec(truth, pred, event_level = "second")
      })
    ) %>%
    slice_max(f1, n = 1, with_ties = FALSE)
}

eval_with_f1_thr <- function(train_truth, train_prob, test_truth, test_prob) {
  thr <- best_f1_threshold(train_truth, train_prob)$threshold[[1]]

  test_truth <- factor(test_truth, levels = c("no", "yes"))
  test_class <- factor(dplyr::if_else(test_prob >= thr, "yes", "no"),
                       levels = c("no", "yes"))

  tibble(
    Accuracy  = yardstick::accuracy_vec(test_truth, test_class),
    F1_Score  = yardstick::f_meas_vec(test_truth, test_class, event_level = "second"),
    AUC       = yardstick::roc_auc_vec(test_truth, test_prob, event_level = "second"),
    Threshold = thr
  )
}

get_prob_yes <- function(fit_obj, new_df) {
  predict(fit_obj, new_df, type = "prob")$.pred_yes
}

eval_model <- function(name, fit_obj, train_df, test_df) {
  train_prob <- get_prob_yes(fit_obj, train_df)
  test_prob  <- get_prob_yes(fit_obj, test_df)

  eval_with_f1_thr(train_df$romantic, train_prob, test_df$romantic, test_prob) %>%
    mutate(Model = name) %>%
    dplyr::select(Model, Accuracy, F1_Score, AUC, Threshold)
}

eval_avg_prob <- function(name, fit_list, train_df, test_df) {
  train_prob_mat <- sapply(fit_list, function(fit) get_prob_yes(fit, train_df))
  test_prob_mat  <- sapply(fit_list, function(fit) get_prob_yes(fit, test_df))

  train_prob <- rowMeans(train_prob_mat)
  test_prob  <- rowMeans(test_prob_mat)

  eval_with_f1_thr(train_df$romantic, train_prob, test_df$romantic, test_prob) %>%
    mutate(Model = name) %>%
    dplyr::select(Model, Accuracy, F1_Score, AUC, Threshold)
}

results_all <- bind_rows(
  eval_model("Logistic (baseline)", logit_fit, train_data_mod, test_data_mod),
  eval_model("Random Forest (tuned)", rf_fit, train_data_mod, test_data_mod),
  eval_model("Decision Tree (tuned)", tree_fit, train_data_mod, test_data_mod),
  eval_model("Elastic Net", enet_fit, train_data_mod, test_data_mod),
  eval_model("SVM RBF", svm_fit, train_data_mod, test_data_mod),
  eval_model("kNN", knn_fit, train_data_mod, test_data_mod),
  eval_model("Neural Net (MLP)", mlp_fit, train_data_mod, test_data_mod),
  eval_model("XGBoost", xgb_fit, train_data_mod, test_data_mod),
  eval_avg_prob("Avg Prob (RF+XGB+SVM)", list(rf_fit, xgb_fit, svm_fit), train_data_mod, test_data_mod),
  eval_model("Stacking Ensemble", stack_fit_all, train_data_mod, test_data_mod)
)

results_F <- bind_rows(
  eval_model("Logistic (baseline)", logit_fit_F, train_F_mod, test_F_mod),
  eval_model("Decision Tree (tuned)", logit_fit_F, train_F_mod, test_F_mod),
  eval_model("Random Forest (tuned)", rf_fit_F, train_F_mod, test_F_mod),
  eval_model("Elastic Net", enet_fit_F, train_F_mod, test_F_mod),
  eval_model("SVM RBF", svm_fit_F, train_F_mod, test_F_mod),
  eval_model("kNN", knn_fit_F, train_F_mod, test_F_mod),
  eval_model("Neural Net (MLP)", mlp_fit_F, train_F_mod, test_F_mod),
  eval_model("XGBoost", xgb_fit_F, train_F_mod, test_F_mod),
  eval_avg_prob("Avg Prob (RF+XGB+SVM)", list(rf_fit_F, xgb_fit_F, svm_fit_F), train_F_mod, test_F_mod),
  eval_model("Stacking Ensemble", stack_fit_F, train_F_mod, test_F_mod)
)

results_M <- bind_rows(
  eval_model("Logistic (baseline)", logit_fit_M, train_M_mod, test_M_mod),
  eval_model("Decision Tree (tuned)", logit_fit_M, train_M_mod, test_M_mod),
  eval_model("Random Forest (tuned)", rf_fit_M, train_M_mod, test_M_mod),
  eval_model("Elastic Net", enet_fit_M, train_M_mod, test_M_mod),
  eval_model("SVM RBF", svm_fit_M, train_M_mod, test_M_mod),
  eval_model("kNN", knn_fit_M, train_M_mod, test_M_mod),
  eval_model("Neural Net (MLP)", mlp_fit_M, train_M_mod, test_M_mod),
  eval_model("XGBoost", xgb_fit_M, train_M_mod, test_M_mod),
  eval_avg_prob("Avg Prob (RF+XGB+SVM)", list(rf_fit_M, xgb_fit_M, svm_fit_M), train_M_mod, test_M_mod),
  eval_model("Stacking Ensemble", stack_fit_M, train_M_mod, test_M_mod)
)

results_all
results_F
results_M
```

```{r}
get_roc_df <- function(truth, prob, model_name, sex_label) {
  yardstick::roc_curve(tibble(truth = truth, prob = prob), truth, prob, event_level = "second") %>%
    dplyr::mutate(Model = model_name, Sex = sex_label)
}

pred_F_logit <- predict(logit_fit_F, test_F_mod, type = "prob")$.pred_yes
pred_F_rf    <- predict(rf_fit_F,    test_F_mod, type = "prob")$.pred_yes
pred_F_svm   <- predict(svm_fit_F,   test_F_mod, type = "prob")$.pred_yes
pred_F_xgb   <- predict(xgb_fit_F,   test_F_mod, type = "prob")$.pred_yes
pred_F_stack <- predict(stack_fit_F, test_F_mod, type = "prob")$.pred_yes

roc_F <- dplyr::bind_rows(
  get_roc_df(test_F_mod$romantic, pred_F_logit, "Logistic (baseline)", "F"),
  get_roc_df(test_F_mod$romantic, pred_F_rf,    "Random Forest (tuned)", "F"),
  get_roc_df(test_F_mod$romantic, pred_F_svm,   "SVM RBF", "F"),
  get_roc_df(test_F_mod$romantic, pred_F_xgb,   "XGBoost", "F"),
  get_roc_df(test_F_mod$romantic, pred_F_stack, "Stacking", "F")
)

pred_M_logit <- predict(logit_fit_M, test_M_mod, type = "prob")$.pred_yes
pred_M_rf    <- predict(rf_fit_M,    test_M_mod, type = "prob")$.pred_yes
pred_M_svm   <- predict(svm_fit_M,   test_M_mod, type = "prob")$.pred_yes
pred_M_xgb   <- predict(xgb_fit_M,   test_M_mod, type = "prob")$.pred_yes
pred_M_stack <- predict(stack_fit_M, test_M_mod, type = "prob")$.pred_yes

roc_M <- dplyr::bind_rows(
  get_roc_df(test_M_mod$romantic, pred_M_logit, "Logistic (baseline)", "M"),
  get_roc_df(test_M_mod$romantic, pred_M_rf,    "Random Forest (tuned)", "M"),
  get_roc_df(test_M_mod$romantic, pred_M_svm,   "SVM RBF", "M"),
  get_roc_df(test_M_mod$romantic, pred_M_xgb,   "XGBoost", "M"),
  get_roc_df(test_M_mod$romantic, pred_M_stack, "Stacking", "M")
)

roc_all <- dplyr::bind_rows(roc_F, roc_M)

ggplot(roc_all, aes(x = 1 - specificity, y = sensitivity, color = Model)) +
  geom_path(linewidth = 1.2) +
  geom_abline(linetype = "dashed", color = "gray") +
  facet_wrap(~ Sex) +
  labs(
    title = "ROC Curves by Sex (Top models + Stacking)",
    x = "1 - Specificity",
    y = "Sensitivity",
    color = "Model"
  ) +
  theme_minimal()

```

```{r}
# ===== kNN deep dive (what's going on?) =====
ens_all$knn_best
ens_F$knn_best
ens_M$knn_best

# kNN tuning curve (All)
knn_auc_curve <- ens_all$knn_res %>%
  collect_metrics() %>%
  dplyr::filter(.metric == "roc_auc")

ggplot(knn_auc_curve, aes(x = neighbors, y = mean, color = weight_func)) +
  geom_line() +
  geom_point(alpha = 0.6) +
  labs(
    title = "kNN tuning curve (All): CV ROC AUC vs k",
    x = "Number of neighbors (k)",
    y = "CV ROC AUC",
    color = "weight_func"
  ) +
  theme_minimal()
```

```{r}
students %>%
  group_by(sex, romantic) %>%
  summarise(
    freetime_pct  = mean(freetime_pct, na.rm = TRUE),
    goout_pct     = mean(goout_pct, na.rm = TRUE),
    studytime_pct = mean(studytime_pct, na.rm = TRUE),
    alc_mean_pct  = mean(alc_mean_pct, na.rm = TRUE),
    Medu_pct = mean(Medu_pct, na.rm = TRUE),
Fedu_pct = mean(Fedu_pct, na.rm = TRUE),
edu_gap_pct = mean(edu_gap_pct, na.rm = TRUE),
    .groups = "drop"
  )
```
