---
title: "Predict Student Romantic Relationship Status"
author: "Weiyang Sheng"
format: pdf
editor: visual
---

```{r}
#| echo: false
#| message: false
#| warning: false

## keep this chunk!
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  tidy = "styler",
  fig.align = "center"
)
```

## Project setup

This chunk loads the packages used throughout the analysis and fixes the random seed so the workflow is reproducible. It provides the software foundation for the entire project.

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidymodels)
library(vip)
library(xgboost)
library(ranger)
library(broom)

options(yardstick.event_first = FALSE)
set.seed(123)
```

## Data import and sample construction

This chunk reads the Math and Portuguese course datasets, identifies overlapping students, and constructs a cleaner combined sample. Its role in the paper is to show that the final analysis is based on a more careful student-level dataset rather than a simple row bind of the two files.

```{r import}
student_por <- read_csv("student-por.csv")
student_mat <- read_csv("student-mat.csv")

student_mat <- student_mat %>% mutate(course = "Math")
student_por <- student_por %>% mutate(course = "Portuguese")

id_vars <- c(
  "school","sex","age","address","famsize","Pstatus",
  "Medu","Fedu","Mjob","Fjob","reason","nursery","internet"
)


dup_students <- inner_join(
  student_mat %>% dplyr::select(all_of(id_vars)) %>% distinct(),
  student_por %>% dplyr::select(all_of(id_vars)) %>% distinct(),
  by = id_vars
)

nrow(dup_students)   # overlap count based on id_vars

mode_value <- function(x) {
  x <- x[!is.na(x)]
  if (length(x) == 0) return(NA_character_)
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

rounded_mean_or_na <- function(x) {
  if (all(is.na(x))) return(NA_real_)
  round(mean(as.numeric(x), na.rm = TRUE))
}

mean_or_na <- function(x) {
  if (all(is.na(x))) return(NA_real_)
  mean(as.numeric(x), na.rm = TRUE)
}

# 这些变量保持原始离散等级，避免合并后出现 1.5、2.5 这种奇怪水平
rounded_merge_vars <- c(
  "traveltime", "studytime", "failures", "famrel",
  "freetime", "goout", "Dalc", "Walc", "health"
)

# 这些变量允许用均值合并
mean_merge_vars <- c("absences", "G1", "G2", "G3")

# yes/no 变量用众数合并
yesno_merge_vars <- c("schoolsup", "famsup", "paid", "activities", "higher", "romantic")

# 其他分类变量用众数合并
categorical_merge_vars <- c("guardian")

student_all <- bind_rows(student_mat, student_por) %>%
  tidyr::unite("student_id", all_of(id_vars), sep = "__", remove = FALSE) %>%
  dplyr::group_by(student_id) %>%
  dplyr::summarise(
    dplyr::across(all_of(id_vars), ~ dplyr::first(.x)),
    course = dplyr::if_else(dplyr::n_distinct(course) > 1, "Both", dplyr::first(course)),
    dplyr::across(any_of(rounded_merge_vars), rounded_mean_or_na),
    dplyr::across(any_of(mean_merge_vars), mean_or_na),
    dplyr::across(any_of(yesno_merge_vars), ~ mode_value(as.character(.x))),
    dplyr::across(any_of(categorical_merge_vars), ~ mode_value(as.character(.x))),
    .groups = "drop"
  ) %>%
  dplyr::select(-student_id)
nrow(student_all)
readr::write_csv(student_all, "student_all_merged.csv")

```

```{r}
getwd()
```

## Data cleaning and feature engineering

This chunk converts raw variables into analysis-ready predictors, including parental education percentages, SES-style job groupings, age groups, and alcohol-related measures. It also creates the train/test split and the sex-specific subsets, which are central to the predictive and moderator-based design of the paper.

```{r}
students <- student_all %>%
  mutate(
    romantic = factor(romantic, levels = c("no", "yes")),
    sex = factor(sex),
    course = factor(course, levels = c("Math", "Portuguese", "Both")),
    school = factor(school),
    address = factor(address),
    famsize = factor(famsize),
    Pstatus = factor(Pstatus),
    Mjob = factor(Mjob),
    Fjob = factor(Fjob),
    guardian = factor(guardian),

    activities = factor(activities, levels = c("no", "yes")),
    higher = factor(higher, levels = c("no", "yes")),
    internet = factor(internet, levels = c("no", "yes")),

    Medu = as.numeric(Medu),
    Fedu = as.numeric(Fedu),
    age = as.numeric(age),
    traveltime = as.numeric(traveltime),
    studytime = as.numeric(studytime),
    failures = as.numeric(failures),
    famrel = as.numeric(famrel),
    freetime = as.numeric(freetime),
    goout = as.numeric(goout),
    Dalc = as.numeric(Dalc),
    Walc = as.numeric(Walc),
    health = as.numeric(health),
    absences = as.numeric(absences),
    G1 = as.numeric(G1),
    G2 = as.numeric(G2),
    G3 = as.numeric(G3),

    Medu_pct = 100 * Medu / 4,
    Fedu_pct = 100 * Fedu / 4,
    edu_gap_pct = abs(Medu_pct - Fedu_pct),

    Mjob_ses = case_when(
      Mjob %in% c("teacher", "health") ~ "higher",
      Mjob %in% c("services") ~ "middle",
      TRUE ~ "lower"
    ),
    Mjob_ses = factor(Mjob_ses, levels = c("lower", "middle", "higher")),

    Fjob_ses = case_when(
      Fjob %in% c("teacher", "health") ~ "higher",
      Fjob %in% c("services") ~ "middle",
      TRUE ~ "lower"
    ),
    Fjob_ses = factor(Fjob_ses, levels = c("lower", "middle", "higher")),

    age_group = case_when(
      age <= 16 ~ "15-16",
      age == 17 ~ "17",
      age == 18 ~ "18",
      age >= 19 ~ "19+"
    ),
    age_group = factor(age_group, levels = c("15-16", "17", "18", "19+")),

    alc_mean = (Dalc + Walc) / 2,
    alc_risk = factor(
      if_else(Walc >= 4 | Dalc >= 3, "higher", "lower"),
      levels = c("lower", "higher")
    ),

    freetime_pct = 100 * (freetime - 1) / 4,
    goout_pct = 100 * (goout - 1) / 4,
    studytime_pct = 100 * (studytime - 1) / 3,
    alc_mean_pct = 100 * (alc_mean - 1) / 4
  )

set.seed(123)
splits <- initial_split(students, prop = 0.8, strata = romantic)

train_data <- training(splits)
test_data  <- testing(splits)

train_data_mod <- train_data
test_data_mod  <- test_data

train_F <- train_data %>% filter(sex == "F")
test_F  <- test_data  %>% filter(sex == "F")

train_M <- train_data %>% filter(sex == "M")
test_M  <- test_data  %>% filter(sex == "M")

train_F_mod <- train_F %>% dplyr::select(-sex)
test_F_mod  <- test_F  %>% dplyr::select(-sex)

train_M_mod <- train_M %>% dplyr::select(-sex)
test_M_mod  <- test_M  %>% dplyr::select(-sex)

prop.table(table(train_data$romantic))
prop.table(table(test_data$romantic))
```

## Plot analysis: mother’s education

This chunk explores whether mother’s education is associated with relationship status after centering the pattern around each sex-specific baseline. Its role is to test whether maternal education carries a strong or weak descriptive signal before modeling begins. \## Plot analysis: father’s education

This chunk repeats the same descriptive analysis for father’s education. Its role is to compare maternal and paternal education effects and to see whether either side of family education separates the outcome more clearly.

```{r}
edu_base <- students %>%
  group_by(sex) %>%
  summarise(base = mean(romantic == "yes"), .groups = "drop")

edu_parent <- students %>%
  dplyr::select(sex, romantic, Medu_pct, Fedu_pct) %>%
  pivot_longer(
    cols = c(Medu_pct, Fedu_pct),
    names_to = "parent",
    values_to = "edu_pct"
  ) %>%
  mutate(
    parent = recode(parent, Medu_pct = "Mother", Fedu_pct = "Father"),
    edu_pct = factor(edu_pct, levels = c(0, 25, 50, 75, 100))
  ) %>%
  group_by(sex, parent, edu_pct) %>%
  summarise(prop_yes = mean(romantic == "yes"), n = n(), .groups = "drop") %>%
  left_join(edu_base, by = "sex") %>%
  mutate(delta_pp = 100 * (prop_yes - base))

pad_parent <- 2

edu_parent_mother <- edu_parent %>% filter(parent == "Mother")
ylim_m <- range(edu_parent_mother$delta_pp, na.rm = TRUE)

ggplot(edu_parent_mother, aes(x = edu_pct, y = delta_pp, group = 1)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_line(linewidth = 1.2) +
  geom_point(aes(size = n), alpha = 0.9) +
  facet_wrap(~ sex, nrow = 1) +
  scale_y_continuous(labels = function(x) paste0(x, " pp")) +
  coord_cartesian(ylim = c(ylim_m[1] - pad_parent, ylim_m[2] + pad_parent)) +
  labs(
    title = "Mother education and relationship rate (difference from sex average)",
    x = "Education level (0–100%)",
    y = "Difference from sex average (percentage points)",
    size = "N"
  ) +
  theme_minimal()

edu_parent_father <- edu_parent %>% filter(parent == "Father")
ylim_f <- range(edu_parent_father$delta_pp, na.rm = TRUE)

ggplot(edu_parent_father, aes(x = edu_pct, y = delta_pp, group = 1)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_line(linewidth = 1.2) +
  geom_point(aes(size = n), alpha = 0.9) +
  facet_wrap(~ sex, nrow = 1) +
  scale_y_continuous(labels = function(x) paste0(x, " pp")) +
  coord_cartesian(ylim = c(ylim_f[1] - pad_parent, ylim_f[2] + pad_parent)) +
  labs(
    title = "Father education and relationship rate (difference from sex average)",
    x = "Education level (0–100%)",
    y = "Difference from sex average (percentage points)",
    size = "N"
  ) +
  theme_minimal()

```

## Dependency check: family background and going out

This chunk tests whether parental education and parental job SES are related to students’ going-out behavior. Its role in the paper is to examine whether family-background variables may work indirectly through more immediate social-exposure variables.

```{r}
# ===== Dependency check example: does parental SES predict goout? =====
goout_lm <- lm(goout ~ Medu_pct + Fedu_pct + Mjob_ses + Fjob_ses + sex, data = students)
summary(goout_lm)

# A simple visualization (mean goout by parental SES)
students %>%
  group_by(sex, Mjob_ses) %>%
  summarise(goout_mean = mean(goout, na.rm = TRUE), n = n(), .groups = "drop") %>%
  ggplot(aes(x = Mjob_ses, y = goout_mean)) +
  geom_col() +
  facet_wrap(~ sex) +
  labs(
    title = "Is going-out behavior associated with mother's job SES?",
    x = "Mother job SES",
    y = "Mean goout (1–5)"
  ) +
  theme_minimal()
```

## Plot analysis: education gap

This chunk examines whether the gap between mother’s and father’s education levels is associated with relationship status, and whether the pattern differs by sex. Its role is to evaluate whether educational mismatch in the family adds information beyond the two separate parent-education measures.

```{r}
students %>%
  group_by(sex, edu_gap_pct) %>%
  summarise(prop_yes = mean(romantic == "yes"), .groups = "drop") %>%
  ggplot(aes(x = edu_gap_pct, y = prop_yes, color = sex, group = sex)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 2.5) +
  scale_x_continuous(breaks = c(0, 25, 50, 75, 100)) +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  labs(
    title = "Education Gap (0–100%) vs Relationship Rate",
    x = "Gap between parents' education (0–100%)",
    y = "Proportion (romantic = yes)",
    color = "Sex"
  ) +
  theme_minimal()
```

## Group-size check for education gap

This chunk reports the sample size behind each education-gap category. Its role is to prevent over-interpretation of visually extreme points that may be driven by very small groups.

```{r}
students %>%
  group_by(sex, edu_gap_pct) %>%
  summarise(
    n = n(),
    yes_n = sum(romantic == "yes"),
    prop_yes = mean(romantic == "yes"),
    .groups = "drop"
  ) %>%
  arrange(sex, edu_gap_pct)
```

## Plot analysis: education gap relative to each sex baseline

This chunk re-expresses the education-gap pattern as a difference from the average relationship rate within each sex. Its role is to make the education-gap signal easier to interpret in relative rather than raw-percentage terms.

```{r}
gap_base <- students %>%
  group_by(sex) %>%
  summarise(base = mean(romantic == "yes"), .groups = "drop")

gap_df <- students %>%
  mutate(
    gap_group = case_when(
      edu_gap_pct == 0 ~ "0",
      edu_gap_pct == 25 ~ "25",
      edu_gap_pct == 50 ~ "50",
      edu_gap_pct >= 75 ~ "75-100"
    ),
    gap_group = factor(gap_group, levels = c("0", "25", "50", "75-100"))
  ) %>%
  group_by(sex, gap_group) %>%
  summarise(prop_yes = mean(romantic == "yes"), n = n(), .groups = "drop") %>%
  left_join(gap_base, by = "sex") %>%
  mutate(delta_pp = 100 * (prop_yes - base))

ylim_gap <- range(gap_df$delta_pp, na.rm = TRUE)
pad_gap <- 2

ggplot(gap_df, aes(x = gap_group, y = delta_pp, group = sex, color = sex)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_line(linewidth = 1.2) +
  geom_point(aes(size = n), alpha = 0.9) +
  scale_y_continuous(labels = function(x) paste0(x, " pp")) +
  coord_cartesian(ylim = c(ylim_gap[1] - pad_gap, ylim_gap[2] + pad_gap)) +
  labs(
    title = "Education gap and relationship rate (difference from sex average)",
    x = "Gap between parents' education",
    y = "Difference from sex average (percentage points)",
    color = "Sex",
    size = "N"
  ) +
  theme_minimal()
```

## Heatmap analysis: joint parent-education patterns

These heatmaps show total sample size, relationship count, and relationship rate across combinations of mother’s and father’s education levels. Their role is to distinguish between cells that are large and cells that are actually higher in relationship prevalence, while also flagging low-N combinations.

```{r}
edu_combo <- students %>%
  mutate(
    Medu_cat = factor(Medu, levels = 0:4),
    Fedu_cat = factor(Fedu, levels = 0:4)
  ) %>%
  group_by(sex, Medu_cat, Fedu_cat) %>%
  summarise(
    yes_n  = sum(romantic == "yes"),
    no_n   = sum(romantic == "no"),
    n      = n(),
    prop_yes = yes_n / n,
    .groups = "drop"
  ) %>%
  mutate(
    # 小样本标记
    n_flag = if_else(n < 5, "Low N", "OK"),
    
    # rate heatmap: 小样本格子先不显示比例标签
    rate_label = if_else(
      n >= 5,
      paste0(scales::percent(prop_yes, accuracy = 1), "\n(n=", n, ")"),
      paste0("n=", n)
    ),
    
    count_label = paste0("yes=", yes_n, "\nN=", n),
    all_label   = paste0("N=", n),
    
    # 让深色格子用白字，浅色格子用黑字
    text_col_count = if_else(yes_n >= median(yes_n, na.rm = TRUE), "white", "black"),
    text_col_all   = if_else(n >= median(n, na.rm = TRUE), "white", "black"),
    text_col_rate  = if_else(prop_yes >= 0.50, "white", "black"),
    
    # 给比例图用：低N直接置为NA，显示成灰色
    prop_yes_plot = if_else(n >= 5, prop_yes, NA_real_)
  )

fill_max_n   <- max(edu_combo$n, na.rm = TRUE)
fill_max_yes <- max(edu_combo$yes_n, na.rm = TRUE)

plot_heatmap_all_n <- function(df_sex, sex_label) {
  ggplot(df_sex, aes(x = Medu_cat, y = Fedu_cat, fill = n)) +
    geom_tile(color = "white", linewidth = 0.5) +
    geom_text(aes(label = all_label, color = text_col_all), size = 3) +
    scale_color_identity() +
    scale_fill_gradient(
      low = "white",
      high = "steelblue",
      limits = c(0, fill_max_n),
      breaks = scales::pretty_breaks(n = 5),
      na.value = "grey90"
    ) +
    labs(
      title = paste0("All students by parents' education (", sex_label, ")"),
      x = "Mother education (0–4)",
      y = "Father education (0–4)",
      fill = "Total N"
    ) +
    theme_minimal()
}

plot_heatmap_yes_n <- function(df_sex, sex_label) {
  ggplot(df_sex, aes(x = Medu_cat, y = Fedu_cat, fill = yes_n)) +
    geom_tile(color = "white", linewidth = 0.5) +
    geom_text(aes(label = count_label, color = text_col_count), size = 3) +
    scale_color_identity() +
    scale_fill_gradient(
      low = "white",
      high = "orange",
      limits = c(0, fill_max_yes),
      breaks = scales::pretty_breaks(n = 5),
      na.value = "grey90"
    ) +
    labs(
      title = paste0("Students in a relationship by parents' education (", sex_label, ")"),
      x = "Mother education (0–4)",
      y = "Father education (0–4)",
      fill = "Yes count"
    ) +
    theme_minimal()
}

plot_heatmap_yes_rate <- function(df_sex, sex_label) {
  ggplot(df_sex, aes(x = Medu_cat, y = Fedu_cat, fill = prop_yes_plot)) +
    geom_tile(color = "white", linewidth = 0.5) +
    geom_text(aes(label = rate_label, color = text_col_rate), size = 3) +
    scale_color_identity() +
    scale_fill_gradient2(
      low = "blue",
      mid = "white",
      high = "red",
      midpoint = 0.50,
      limits = c(0, 1),
      breaks = seq(0, 1, by = 0.25),
      labels = scales::percent_format(accuracy = 1),
      na.value = "grey85"
    ) +
    labs(
      title = paste0("Relationship rate by parents' education (", sex_label, ", grey = low N < 5)"),
      x = "Mother education (0–4)",
      y = "Father education (0–4)",
      fill = "Yes rate"
    ) +
    theme_minimal()
}

# Female
plot_heatmap_all_n(edu_combo %>% filter(sex == "F"), "Female")
plot_heatmap_yes_n(edu_combo %>% filter(sex == "F"), "Female")
plot_heatmap_yes_rate(edu_combo %>% filter(sex == "F"), "Female")

# Male
plot_heatmap_all_n(edu_combo %>% filter(sex == "M"), "Male")
plot_heatmap_yes_n(edu_combo %>% filter(sex == "M"), "Male")
plot_heatmap_yes_rate(edu_combo %>% filter(sex == "M"), "Male")
```

## Basic descriptive comparison by sex

This chunk gives the overall relationship rate for female and male students. Its role is to establish the baseline sex difference before moving into subgroup-specific modeling.

```{r}
students %>%
  group_by(sex) %>%
  summarise(prop_yes = mean(romantic == "yes"), n = n(), .groups = "drop") %>%
  ggplot(aes(x = sex, y = prop_yes)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  labs(
    title = "Relationship Rate by Sex",
    x = "Sex",
    y = "Proportion (romantic = yes)"
  ) +
  theme_minimal()

```

## Plot analysis: parental job SES

This chunk explores whether broader mother/father job SES groupings relate to relationship status. Its role is to test whether a simplified family-resource representation is descriptively useful before entering the predictive stage.

```{r}
students %>%
  dplyr::select(sex, romantic, Mjob_ses, Fjob_ses) %>%
  pivot_longer(cols = c(Mjob_ses, Fjob_ses), names_to = "parent", values_to = "job_ses") %>%
  mutate(parent = recode(parent, Mjob_ses = "Mother", Fjob_ses = "Father")) %>%
  group_by(sex, parent, job_ses) %>%
  summarise(prop_yes = mean(romantic == "yes"), .groups = "drop") %>%
  ggplot(aes(x = job_ses, y = prop_yes)) +
  geom_col() +
  facet_grid(sex ~ parent) +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  labs(
    title = " Parental Job SES vs Relationship Rate (by Sex)",
    x = "Job SES",
    y = "Proportion (romantic = yes)"
  ) +
  theme_minimal()

```

## Plot analysis: free time

This chunk visualizes the relationship between free time and romantic status by sex. Its role is to inspect whether one of the project’s key social-exposure variables shows a descriptive pattern worth carrying into the modeling discussion.

```{r}
students %>%
  group_by(sex, freetime_pct) %>%
  summarise(prop_yes = mean(romantic == "yes"), .groups = "drop") %>%
  ggplot(aes(x = freetime_pct, y = prop_yes, color = sex, group = sex)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3) +
  scale_x_continuous(breaks = c(0, 25, 50, 75, 100)) +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  labs(
    title = "Free-time (0–100%) × Sex",
    x = "Free-time (0–100%)",
    y = "Proportion (romantic = yes)",
    color = "Sex"
  ) +
  theme_minimal()
```

## Plot analysis: alcohol risk

This chunk compares relationship prevalence across lower- and higher-risk alcohol groups for female and male students. Its role is to test whether alcohol-related social behavior is descriptively linked to romantic status.

```{r}
students %>%
  group_by(sex, alc_risk) %>%
  summarise(prop_yes = mean(romantic == "yes"), .groups = "drop") %>%
  ggplot(aes(x = alc_risk, y = prop_yes)) +
  geom_col() +
  facet_wrap(~ sex) +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  labs(
    title = "Alcohol Risk vs Relationship Rate (by Sex)",
    x = "Alcohol risk",
    y = "Proportion (romantic = yes)"
  ) +
  theme_minimal()
```

## Plot analysis: adjusted absences effect

This chunk uses a logistic model to estimate the adjusted relationship between absences and the probability of being in a relationship. Its role is to move beyond raw grouped percentages and show a smoother, more interpretable pattern for one of the clearest predictors in the current results.

```{r}
# ===== Adjusted prediction plot: Absences =====
students_abs <- students %>%
  mutate(abs_cap = pmin(absences, 15))   # 把极端高缺课值压到 15，避免尾部太稀疏

abs_vis_fit <- glm(
  I(romantic == "yes") ~ poly(abs_cap, 2, raw = TRUE) * sex,
  family = binomial(),
  data = students_abs
)

abs_grid <- expand.grid(
  abs_cap = 0:15,
  sex = levels(students_abs$sex)
)

abs_pred <- predict(abs_vis_fit, newdata = abs_grid, type = "link", se.fit = TRUE)

abs_plot_df <- abs_grid %>%
  mutate(
    fit_link = abs_pred$fit,
    se_link  = abs_pred$se.fit,
    prob  = plogis(fit_link),
    lower = plogis(fit_link - 1.96 * se_link),
    upper = plogis(fit_link + 1.96 * se_link)
  )

ggplot(abs_plot_df, aes(x = abs_cap, y = prob, color = sex, fill = sex)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.18, color = NA) +
  geom_line(linewidth = 1.2) +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  labs(
    title = "Adjusted relationship probability by absences",
    subtitle = "Curves are based on a logistic model with a quadratic term; absences above 15 are capped",
    x = "Absences (capped at 15)",
    y = "Predicted probability of being in a relationship",
    color = "Sex",
    fill = "Sex"
  ) +
  theme_minimal()
```

## Plot analysis: internet access

This chunk visualizes internet access as a simple and interpretable binary predictor, with labels showing the sex-specific percentage-point difference from the no-internet group. Its role is to highlight a direct opportunity/resource variable that is easier to interpret than the weaker parent-education effects.

```{r}
# ===== Internet: clearer percentage-point comparison =====
internet_df <- students %>%
  group_by(sex, internet) %>%
  summarise(
    prop_yes = mean(romantic == "yes"),
    n = n(),
    .groups = "drop"
  ) %>%
  group_by(sex) %>%
  mutate(
    base_no = prop_yes[internet == "no"],
    diff_pp = 100 * (prop_yes - base_no)
  ) %>%
  ungroup()

ggplot(internet_df, aes(x = internet, y = prop_yes, fill = sex)) +
  geom_col(position = position_dodge(width = 0.75), width = 0.65) +
  geom_text(
    aes(
      label = if_else(
        internet == "yes",
        paste0(scales::percent(prop_yes, accuracy = 1), "\n", sprintf("%+.0f pp", diff_pp), "\nN=", n),
        paste0(scales::percent(prop_yes, accuracy = 1), "\nBaseline\nN=", n)
      )
    ),
    position = position_dodge(width = 0.75),
    vjust = -0.2,
    size = 3
  ) +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  labs(
    title = "Internet access and relationship rate",
    subtitle = "Labels show the sex-specific percentage-point difference from the no-internet group",
    x = "Internet at home",
    y = "Proportion in relationship (romantic = yes)",
    fill = "Sex"
  ) +
  theme_minimal()
```

## Plot analysis: grouped absences

This chunk provides a simpler grouped view of absences and relationship prevalence across the full sample. Its role is mainly descriptive, while the adjusted absences plot provides the stronger interpretation.

```{r}
abs_df_all <- students %>%
  mutate(
    abs_group = case_when(
      absences == 0 ~ "0",
      absences <= 2 ~ "1-2",
      absences <= 5 ~ "3-5",
      absences <= 10 ~ "6-10",
      TRUE ~ "11+"
    ),
    abs_group = factor(abs_group, levels = c("0", "1-2", "3-5", "6-10", "11+"))
  ) %>%
  group_by(abs_group) %>%
  summarise(
    prop_yes = mean(romantic == "yes"),
    n = n(),
    .groups = "drop"
  )

ggplot(abs_df_all, aes(x = abs_group, y = prop_yes)) +
  geom_col() +
  geom_text(aes(label = paste0("n=", n)), vjust = -0.3, size = 3) +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  labs(
    title = "Relationship rate by absence group (All students)",
    x = "Absence group",
    y = "Proportion in relationship (romantic = yes)"
  ) +
  theme_minimal()
```

## Plot analysis: adjusted study time effect

This chunk shows the model-adjusted probability of being in a relationship across study-time levels, separately by sex. Its role is to present study time as a level-based predictor rather than forcing it into a misleading straight-line interpretation.

```{r}
# ===== Adjusted prediction plot: Study time =====
study_vis_fit <- glm(
  I(romantic == "yes") ~ factor(studytime) * sex,
  family = binomial(),
  data = students
)

study_grid <- expand.grid(
  studytime = sort(unique(students$studytime)),
  sex = levels(students$sex)
)

study_pred <- predict(study_vis_fit, newdata = study_grid, type = "link", se.fit = TRUE)

study_plot_df <- study_grid %>%
  mutate(
    fit_link = study_pred$fit,
    se_link  = study_pred$se.fit,
    prob  = plogis(fit_link),
    lower = plogis(fit_link - 1.96 * se_link),
    upper = plogis(fit_link + 1.96 * se_link)
  ) %>%
  left_join(
    students %>%
      group_by(sex, studytime) %>%
      summarise(n = n(), .groups = "drop"),
    by = c("sex", "studytime")
  )

ggplot(study_plot_df, aes(x = factor(studytime), y = prob, fill = sex)) +
  geom_col(position = position_dodge(width = 0.75), width = 0.65) +
  geom_errorbar(
    aes(ymin = lower, ymax = upper),
    position = position_dodge(width = 0.75),
    width = 0.15
  ) +
  geom_text(
    aes(label = paste0("N=", n)),
    position = position_dodge(width = 0.75),
    vjust = -0.3,
    size = 3
  ) +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  labs(
    title = "Adjusted relationship probability by study time",
    subtitle = "Bars show model-adjusted predicted probability; error bars show 95% CI",
    x = "Study time level",
    y = "Predicted probability of being in a relationship",
    fill = "Sex"
  ) +
  theme_minimal()
```

## Modeling approach: preprocessing recipe

This chunk defines the preprocessing pipeline used before machine-learning models are fitted. Its role is to make the predictive comparison fair by standardizing numeric variables, collapsing rare categories, removing redundant predictors, and creating dummy-coded inputs where needed.

```{r}
make_recipe <- function(df) {
  recipe(romantic ~ ., data = df) %>%
    step_rm(
      course, age, Dalc, Walc, Mjob, Fjob, Medu, Fedu,
      freetime_pct, goout_pct, studytime_pct, alc_mean_pct
    ) %>%
    step_other(all_nominal_predictors(), threshold = 0.03, other = "other_collapsed") %>%
    step_nzv(all_predictors()) %>%
    step_normalize(all_numeric_predictors()) %>%
    step_dummy(all_nominal_predictors()) %>%
    step_zv(all_predictors())
}

rec_all <- make_recipe(train_data_mod)
rec_F   <- make_recipe(train_F_mod)
rec_M   <- make_recipe(train_M_mod)
```

## Baseline model: logistic regression

This chunk fits logistic regression models for the pooled sample and the two sex-specific subsets. Its role is to provide an interpretable statistical baseline before comparing more flexible nonlinear models.

```{r}
logit_spec <- logistic_reg() %>% set_engine("glm")

logit_wf <- workflow() %>%
  add_recipe(rec_all) %>%
  add_model(logit_spec)
logit_fit <- fit(logit_wf, data = train_data_mod)

logit_wf_F <- workflow() %>%
  add_recipe(rec_F) %>%
  add_model(logit_spec)
logit_fit_F <- fit(logit_wf_F, data = train_F_mod)

logit_wf_M <- workflow() %>%
  add_recipe(rec_M) %>%
  add_model(logit_spec)
logit_fit_M <- fit(logit_wf_M, data = train_M_mod)

logit_fit %>%
  extract_fit_parsnip() %>%
  tidy() %>%
  arrange(p.value)

logit_fit_F %>%
  extract_fit_parsnip() %>%
  tidy() %>%
  arrange(p.value)

logit_fit_M %>%
  extract_fit_parsnip() %>%
  tidy() %>%
  arrange(p.value)
```

## Tuned machine-learning models

This chunk tunes two flexible nonlinear models, Random Forest and XGBoost, and compares them against the logistic baseline. Its role is to keep the workflow faster, simpler, and accuracy-first while avoiding unnecessary ensemble complexity.

```{r}
metric_cls <- metric_set(accuracy, roc_auc)

rf_spec_tune <- rand_forest(
  mode = "classification",
  trees = 1000,
  mtry = tune(),
  min_n = tune()
) %>%
  set_engine("ranger", importance = "impurity")

xgb_spec_tune <- boost_tree(
  mode = "classification",
  trees = 800,
  tree_depth = tune(),
  learn_rate = tune(),
  mtry = tune(),
  min_n = tune(),
  loss_reduction = tune(),
  sample_size = tune()
) %>%
  set_engine("xgboost", verbosity = 0)

get_n_pred <- function(train_df, rec_obj) {
  rec_obj %>%
    prep(training = train_df, retain = TRUE) %>%
    juice() %>%
    dplyr::select(-romantic) %>%
    ncol()
}

fit_best <- function(wf_obj, res_obj, train_df_inner) {
  best_params <- tune::select_best(res_obj, metric = "accuracy")
  wf_final <- tune::finalize_workflow(wf_obj, best_params)
  fit(wf_final, data = train_df_inner)
}

fit_simple_models <- function(train_df, rec_obj, v = 5) {
  set.seed(123)
  folds <- vfold_cv(train_df, v = v, strata = romantic)

  ctrl <- tune::control_grid(
    save_workflow = TRUE,
    event_level = "second"
  )

  n_pred <- get_n_pred(train_df, rec_obj)
  max_mtry <- max(1, min(n_pred, 15))

  rf_wf <- workflow() %>%
    add_recipe(rec_obj) %>%
    add_model(rf_spec_tune)

  xgb_wf <- workflow() %>%
    add_recipe(rec_obj) %>%
    add_model(xgb_spec_tune)

  rf_params <- hardhat::extract_parameter_set_dials(rf_spec_tune) %>%
    update(
      mtry = dials::mtry(c(1, max_mtry)),
      min_n = dials::min_n(c(2, 20))
    )

  rf_grid <- dials::grid_space_filling(rf_params, size = 15)

  xgb_params <- hardhat::extract_parameter_set_dials(xgb_spec_tune) %>%
    update(
      mtry = dials::mtry(c(1, max_mtry)),
      tree_depth = dials::tree_depth(c(2, 8)),
      min_n = dials::min_n(c(2, 20)),
      learn_rate = dials::learn_rate(c(-4, -1)),
      loss_reduction = dials::loss_reduction(c(-5, 1)),
      sample_size = dials::sample_prop(c(0.6, 1.0))
    )

  xgb_grid <- dials::grid_space_filling(xgb_params, size = 15)

  set.seed(123)
  rf_res <- tune_grid(
    rf_wf,
    resamples = folds,
    grid = rf_grid,
    metrics = metric_cls,
    control = ctrl
  )

  set.seed(123)
  xgb_res <- tune_grid(
    xgb_wf,
    resamples = folds,
    grid = xgb_grid,
    metrics = metric_cls,
    control = ctrl
  )

  list(
    rf_fit = fit_best(rf_wf, rf_res, train_df),
    xgb_fit = fit_best(xgb_wf, xgb_res, train_df),
    rf_best = tune::select_best(rf_res, metric = "accuracy"),
    xgb_best = tune::select_best(xgb_res, metric = "accuracy"),
    rf_res = rf_res,
    xgb_res = xgb_res
  )
}

mods_all <- fit_simple_models(train_data_mod, rec_all, v = 5)
mods_F   <- fit_simple_models(train_F_mod, rec_F, v = 5)
mods_M   <- fit_simple_models(train_M_mod, rec_M, v = 5)

rf_fit   <- mods_all$rf_fit
xgb_fit  <- mods_all$xgb_fit

rf_fit_F  <- mods_F$rf_fit
xgb_fit_F <- mods_F$xgb_fit

rf_fit_M  <- mods_M$rf_fit
xgb_fit_M <- mods_M$xgb_fit

mods_all$rf_best
mods_all$xgb_best
```

## Model performance comparison

This chunk compares the three retained models on the held-out test set. Its role in the paper is to determine which simplified model set performs best in the pooled sample and in the female and male subgroups, using Accuracy, F1, and AUC together rather than relying on a single metric.

```{r}
safe_f1 <- function(truth, pred) {
  out <- suppressWarnings(
    yardstick::f_meas_vec(truth, pred, event_level = "second")
  )
  if (is.na(out)) 0 else out
}

best_threshold <- function(truth, prob,
                           optimize_for = c("accuracy", "f1"),
                           grid = seq(0.05, 0.95, by = 0.01)) {
  optimize_for <- match.arg(optimize_for)
  truth <- factor(truth, levels = c("no", "yes"))

  tibble(threshold = grid) %>%
    mutate(
      score = purrr::map_dbl(threshold, function(t) {
        pred <- factor(if_else(prob >= t, "yes", "no"),
                       levels = c("no", "yes"))
        if (optimize_for == "accuracy") {
          yardstick::accuracy_vec(truth, pred)
        } else {
          safe_f1(truth, pred)
        }
      })
    ) %>%
    slice_max(score, n = 1, with_ties = FALSE)
}

eval_with_train_threshold <- function(train_truth, train_prob, test_truth, test_prob,
                                      optimize_for = "accuracy") {
  thr <- best_threshold(
    train_truth, train_prob,
    optimize_for = optimize_for
  )$threshold[[1]]

  test_truth <- factor(test_truth, levels = c("no", "yes"))
  test_class <- factor(if_else(test_prob >= thr, "yes", "no"),
                       levels = c("no", "yes"))

  tibble(
    Accuracy = yardstick::accuracy_vec(test_truth, test_class),
    F1_Score = safe_f1(test_truth, test_class),
    AUC = yardstick::roc_auc_vec(test_truth, test_prob, event_level = "second"),
    Threshold = thr,
    OptimizeFor = optimize_for
  )
}

get_prob_yes <- function(fit_obj, new_df) {
  predict(fit_obj, new_data = new_df, type = "prob")$.pred_yes
}

eval_model <- function(name, fit_obj, train_df, test_df,
                       optimize_for = "accuracy") {
  train_prob <- get_prob_yes(fit_obj, train_df)
  test_prob  <- get_prob_yes(fit_obj, test_df)

  eval_with_train_threshold(
    train_df$romantic, train_prob,
    test_df$romantic,  test_prob,
    optimize_for = optimize_for
  ) %>%
    mutate(Model = name) %>%
    dplyr::select(Model, Accuracy, F1_Score, AUC, Threshold, OptimizeFor)
}

single_model_names <- c(
  "Logistic (baseline)",
  "Random Forest (tuned)",
  "XGBoost"
)

results_all <- bind_rows(
  eval_model("Logistic (baseline)", logit_fit, train_data_mod, test_data_mod, optimize_for = "accuracy"),
  eval_model("Random Forest (tuned)", rf_fit, train_data_mod, test_data_mod, optimize_for = "accuracy"),
  eval_model("XGBoost", xgb_fit, train_data_mod, test_data_mod, optimize_for = "accuracy")
) %>%
  arrange(desc(Accuracy), desc(AUC))

results_F <- bind_rows(
  eval_model("Logistic (baseline)", logit_fit_F, train_F_mod, test_F_mod, optimize_for = "accuracy"),
  eval_model("Random Forest (tuned)", rf_fit_F, train_F_mod, test_F_mod, optimize_for = "accuracy"),
  eval_model("XGBoost", xgb_fit_F, train_F_mod, test_F_mod, optimize_for = "accuracy")
) %>%
  arrange(desc(Accuracy), desc(AUC))

results_M <- bind_rows(
  eval_model("Logistic (baseline)", logit_fit_M, train_M_mod, test_M_mod, optimize_for = "accuracy"),
  eval_model("Random Forest (tuned)", rf_fit_M, train_M_mod, test_M_mod, optimize_for = "accuracy"),
  eval_model("XGBoost", xgb_fit_M, train_M_mod, test_M_mod, optimize_for = "accuracy")
) %>%
  arrange(desc(Accuracy), desc(AUC))

results_all
results_F
results_M
```

## Main model comparison results

This section summarizes the strongest single models by test accuracy. Because the exact values may change after preprocessing or tuning updates, the summary below is generated directly from the current result tables.

```{r}
best_all_single <- results_all %>%
  filter(Model %in% single_model_names) %>%
  slice_max(Accuracy, n = 1, with_ties = FALSE)

best_F_single <- results_F %>%
  filter(Model %in% single_model_names) %>%
  slice_max(Accuracy, n = 1, with_ties = FALSE)

best_M_single <- results_M %>%
  filter(Model %in% single_model_names) %>%
  slice_max(Accuracy, n = 1, with_ties = FALSE)

cat(glue::glue("
In the pooled sample, the best single model by test Accuracy is **{best_all_single$Model[[1]]}**, with Accuracy = {round(best_all_single$Accuracy[[1]], 3)}, F1 = {round(best_all_single$F1_Score[[1]], 3)}, and AUC = {round(best_all_single$AUC[[1]], 3)}.

In the female subgroup, the best single model by test Accuracy is **{best_F_single$Model[[1]]}**, with Accuracy = {round(best_F_single$Accuracy[[1]], 3)}, F1 = {round(best_F_single$F1_Score[[1]], 3)}, and AUC = {round(best_F_single$AUC[[1]], 3)}.

In the male subgroup, the best single model by test Accuracy is **{best_M_single$Model[[1]]}**, with Accuracy = {round(best_M_single$Accuracy[[1]], 3)}, F1 = {round(best_M_single$F1_Score[[1]], 3)}, and AUC = {round(best_M_single$AUC[[1]], 3)}.

Taken together, the current version is now accuracy-first and simplified: it keeps sex in the pooled model, merges overlapping students at the student level, compares only Logistic Regression, Random Forest, and XGBoost, selects tuned hyperparameters by Accuracy, and reports the strongest single classifier in each sample.
"))
```

## ROC curve comparison

This chunk plots ROC curves for the strongest available models in the female and male subgroups. Its role is to compare ranking performance visually, while safely skipping stacking curves if a subgroup-specific stacking ensemble does not retain usable members.

```{r}
get_roc_df <- function(truth, prob, model_name, sex_label) {
  yardstick::roc_curve(
    tibble(truth = truth, prob = prob),
    truth, prob,
    event_level = "second"
  ) %>%
    dplyr::mutate(Model = model_name, Sex = sex_label)
}

# ----- Female -----
roc_F <- bind_rows(
  get_roc_df(test_F_mod$romantic, get_prob_yes(logit_fit_F, test_F_mod), "Logistic (baseline)", "F"),
  get_roc_df(test_F_mod$romantic, get_prob_yes(rf_fit_F, test_F_mod), "Random Forest (tuned)", "F"),
  get_roc_df(test_F_mod$romantic, get_prob_yes(xgb_fit_F, test_F_mod), "XGBoost", "F")
)

# ----- Male -----
roc_M <- bind_rows(
  get_roc_df(test_M_mod$romantic, get_prob_yes(logit_fit_M, test_M_mod), "Logistic (baseline)", "M"),
  get_roc_df(test_M_mod$romantic, get_prob_yes(rf_fit_M, test_M_mod), "Random Forest (tuned)", "M"),
  get_roc_df(test_M_mod$romantic, get_prob_yes(xgb_fit_M, test_M_mod), "XGBoost", "M")
)

roc_all <- bind_rows(roc_F, roc_M)

ggplot(roc_all, aes(x = 1 - specificity, y = sensitivity, color = Model)) +
  geom_path(linewidth = 1.2) +
  geom_abline(linetype = "dashed", color = "gray") +
  facet_wrap(~ Sex) +
  labs(
    title = "ROC Curves by Sex",
    x = "1 - Specificity",
    y = "Sensitivity",
    color = "Model"
  ) +
  theme_minimal()
```

## Odds-ratio summary of selected predictors

This chunk summarizes the direction and approximate strength of several key predictors from the logistic model. Its role is to connect the predictive analysis back to interpretable psychological and behavioral variables.

```{r}
or_df <- logit_fit %>%
  extract_fit_parsnip() %>%
  tidy(conf.int = TRUE) %>%
  filter(term %in% c("absences", "studytime", "internet_yes", "Medu_pct", "Fedu_pct")) %>%
  mutate(
    term = recode(
      term,
      absences = "Absences",
      studytime = "Study time",
      internet_yes = "Internet = yes",
      Medu_pct = "Mother education (%)",
      Fedu_pct = "Father education (%)"
    ),
    OR = exp(estimate),
    OR_low = exp(conf.low),
    OR_high = exp(conf.high)
  )

ggplot(or_df, aes(x = OR, y = reorder(term, OR))) +
  geom_point(size = 3) +
  geom_errorbarh(aes(xmin = OR_low, xmax = OR_high), height = 0.15) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "gray50") +
  labs(
    title = "Odds-ratio summary of selected predictors",
    x = "Odds ratio",
    y = NULL
  ) +
  theme_minimal()
```

## Descriptive summary of key group differences

This final chunk compares average behavioral and family-background features across sex and relationship-status groups. Its role is to reconnect the predictive results to more interpretable group-level patterns at the end of the paper.

```{r}
students %>%
  group_by(sex, romantic) %>%
  summarise(
    freetime_pct  = mean(freetime_pct, na.rm = TRUE),
    goout_pct     = mean(goout_pct, na.rm = TRUE),
    studytime_pct = mean(studytime_pct, na.rm = TRUE),
    alc_mean_pct  = mean(alc_mean_pct, na.rm = TRUE),
    Medu_pct = mean(Medu_pct, na.rm = TRUE),
Fedu_pct = mean(Fedu_pct, na.rm = TRUE),
edu_gap_pct = mean(edu_gap_pct, na.rm = TRUE),
    .groups = "drop"
  )
```

## Best-model predictions: pooled sample

This chunk reports the test-set predictions from the strongest pooled-sample single model under the current accuracy-first results. The best model is selected automatically from `results_all`.

```{r}
fit_lookup_all <- setNames(
  list(logit_fit, rf_fit, xgb_fit),
  single_model_names
)

best_all_single <- results_all %>%
  filter(Model %in% single_model_names) %>%
  slice_max(Accuracy, n = 1, with_ties = FALSE)

best_model_name_all <- best_all_single$Model[[1]]
thr_all_best <- best_all_single$Threshold[[1]]
fit_all_best <- fit_lookup_all[[best_model_name_all]]

pred_all_best <- tibble(
  truth = test_data_mod$romantic,
  prob_yes = get_prob_yes(fit_all_best, test_data_mod)
) %>%
  mutate(
    pred_class = factor(
      if_else(prob_yes >= thr_all_best, "yes", "no"),
      levels = c("no", "yes")
    )
  )

cm_all_best <- yardstick::conf_mat(pred_all_best, truth = truth, estimate = pred_class)

pred_all_best %>% slice_head(n = 10)
cm_all_best
```

This table shows the actual class (`truth`), the predicted probability of being in a relationship (`prob_yes`), and the final predicted class (`pred_class`) after applying the selected threshold. The confusion matrix below summarizes how often the model correctly classified `yes` and `no` cases in the held-out pooled test set.\

## Best-model predictions: female subgroup

This chunk reports the test-set predictions from the strongest single-model classifier in the female subgroup. The best model is selected automatically from `results_F` by test accuracy.

```{r}
fit_lookup_F <- setNames(
  list(logit_fit_F, rf_fit_F, xgb_fit_F),
  single_model_names
)

best_F_single <- results_F %>%
  filter(Model %in% single_model_names) %>%
  slice_max(Accuracy, n = 1, with_ties = FALSE)

best_model_name_F <- best_F_single$Model[[1]]
thr_F_best <- best_F_single$Threshold[[1]]
fit_F_best <- fit_lookup_F[[best_model_name_F]]

pred_F_best <- tibble(
  truth = test_F_mod$romantic,
  prob_yes = get_prob_yes(fit_F_best, test_F_mod)
) %>%
  mutate(
    pred_class = factor(
      if_else(prob_yes >= thr_F_best, "yes", "no"),
      levels = c("no", "yes")
    )
  )

cm_F_best <- yardstick::conf_mat(pred_F_best, truth = truth, estimate = pred_class)

pred_F_best %>% slice_head(n = 10)
cm_F_best
```

This output shows how the best female classifier behaves on unseen students. The table makes the prediction process concrete, while the confusion matrix shows whether the model is more likely to miss true `yes` cases or to incorrectly label `no` cases as `yes`.

## Best-model predictions: male subgroup

This chunk reports the test-set predictions from the strongest single-model classifier in the male subgroup. The best model is selected automatically from `results_M` by test accuracy.

```{r}
fit_lookup_M <- setNames(
  list(logit_fit_M, rf_fit_M, xgb_fit_M),
  single_model_names
)

best_M_single <- results_M %>%
  filter(Model %in% single_model_names) %>%
  slice_max(Accuracy, n = 1, with_ties = FALSE)

best_model_name_M <- best_M_single$Model[[1]]
thr_M_best <- best_M_single$Threshold[[1]]
fit_M_best <- fit_lookup_M[[best_model_name_M]]

pred_M_best <- tibble(
  truth = test_M_mod$romantic,
  prob_yes = get_prob_yes(fit_M_best, test_M_mod)
) %>%
  mutate(
    pred_class = factor(
      if_else(prob_yes >= thr_M_best, "yes", "no"),
      levels = c("no", "yes")
    )
  )

cm_M_best <- yardstick::conf_mat(pred_M_best, truth = truth, estimate = pred_class)

pred_M_best %>% slice_head(n = 10)
cm_M_best
```

This output presents the strongest male classifier on the held-out test set. Because the male subgroup tends to be more sensitive to model choice and sample fluctuation, this chunk is useful for showing how the final selected model behaves at the level of actual case classification.

## Interpretation of best-model prediction outputs

This section converts the confusion matrices into short plain-language interpretations so the classification results are easier to read in the paper.

```{r}
library(glue)

get_conf_counts <- function(cm_obj) {
  tab <- cm_obj$table
  list(
    TN = unname(tab["no",  "no"]),
    FN = unname(tab["no",  "yes"]),
    FP = unname(tab["yes", "no"]),
    TP = unname(tab["yes", "yes"])
  )
}

describe_conf <- function(cm_obj, label, model_name) {
  cc <- get_conf_counts(cm_obj)

  err_sentence <- if (cc$FP > cc$FN) {
    "The model makes more false positives than false negatives, which means it is somewhat more willing to predict `yes`."
  } else if (cc$FN > cc$FP) {
    "The model makes more false negatives than false positives, which means it is somewhat more conservative and misses some true `yes` cases."
  } else {
    "The model shows a fairly balanced error pattern between false positives and false negatives."
  }

  glue("
### {label}

For the **{model_name}** model, the confusion matrix contains **{cc$TN} true negatives**, **{cc$TP} true positives**, **{cc$FP} false positives**, and **{cc$FN} false negatives**. {err_sentence} This helps clarify not only how accurate the model is, but also what kind of mistakes it tends to make.
")
}

cat(describe_conf(cm_all_best, "Pooled sample interpretation", best_model_name_all))
cat(describe_conf(cm_F_best, "Female subgroup interpretation", best_model_name_F))
cat(describe_conf(cm_M_best, "Male subgroup interpretation", best_model_name_M))
```
